SAT Competition 2020\tnoteref{title}

### 2021-02-02 paper/main.tex
checking, sect4

\begin{abstract}
The SAT Competitions constitute a well-established series of yearly open international algorithm implementation competitions
focusing on the Boolean satisfiability (or propositional satisfiability, SAT) problem. 
In this article, we provide
a detailed account on the 2020 instantiation of the SAT Competition, including the 
 new competition tracks and benchmark selection procedures, overview of solving strategies implemented in top-performing solvers, 
and a more detailed analysis of the empirical data obtained from running the competition.
%SAT Competition 2020 stands in the tradition of the series of annual competitive events which motivate and assess the progress in SAT solving. 
%This competition was special as it introduced the new cloud track where SAT Solvers that run on hundreds of processors could compete. 
%Another novelty was the application-specific sub-track of the main track, 
%where solvers competed in solving benchmark instances from one specific domain only, and in this year that was the planning domain. 
%We used new tools to select and distribute benchmark instances and their attributes. 
%In this paper we provide a description of the well-known and the new competition tracks and how we organized them. 
%It follows then a detailed analysis of the results and the strategies of the award winning solvers. 
\end{abstract}

@@

The success of SAT would not be without the persistent efforts of the SAT community to further develop the performance and robustness
of SAT solvers. The SAT Competition series, which a history dating back to early 90s, aims to support and provide further incentives 
for maintaining and furthering this progress. Organized yearly as an international open event, the SAT competitions (and their variants
in the form of SAT Races and SAT Challenge) have a consistent track record of receiving tens of solver submissions yearly, submitted
by the community at large for obtaining a snapshot of the current state-of-the-art in practical SAT solving. Alongside participating solvers,
the competition invite through open calls submissions of benchmark instances representing in particular new interesting applications 
scenarios of SAT solvers. Indeed, in addition to evaluating recently developed solvers, an important aspect of the SAT competition series is
to construct on a yearly basis new benchmark sets, consisting of instances from various different application settings, which together with benchmark sets from previous years constitute a standard dataset for use in research papers and SAT solver development.

This article focuses on the 2020 instantiation of the SAT competitions. 
To this end, we provide
a detailed account of SAT Competition 2020 in terms of organizational details, competition tracks,
participating solvers, benchmarks, and the empirical results from the competition.
In terms of competition tracks, two new tracks, namely the cloud track and an application-specific track, where introduced in 2020 
in addition to the already earlier established main, parallel, and incremental tracks; we
provide motivation and the new organizational details for both of the tracks.
In terms of solvers, we provide an overview of solving strategies and other details implemented in top-performing solvers
from the competition, complementing the individual solver descriptions available in the 2020 competition proceedings.
 As for benchmarks, we describe how the 2020 benchmark sets were constructed for each of the competition tracks, with 
an overview of the benchmarks contributed to the 2020 competition.
In terms of empirical results we provide further analysis on the results going beyond the standard rankings provided
on the SAT competition web pages.\footnote{\url{https://satcompetition.github.io/2020/}}. 
Finally, we also provide a discussion on lessons learned and ideas for future editions of SAT competitions.

The rest of this article is organized as follows. 
We start by providing an overview on the competition, including details on and motivations for the several
competition tracks, the rules and other technical requirements of the competition, the ranking schemes used
in evaluating the competing solvers, 
and the computing environments used for executing the competition
(Section~\ref{sec:overview}).
We then provide an overview of the benchmark sets used in evaluating the solvers, including their origins and the selection
process used for constructing the sets (Section~\ref{sec:instances}).
In Section~\ref{sec:results} we provide an overview of the competition results %together with a survey on the
followed by a survey on 
the solving strategies implemented distinctly in top-ranking solvers in Section~\ref{sec:winners}. 
Going considerably beyond the plain competition rankings, in Section~\ref{sec:analysis}, we provide more in-depth analysis of the competition 
data from different perspectives, including correlation analysis of runtime performance of solvers and marginal contributions
of individual solvers to the ``virtual best solver'' and 
portfolios constructed from the competing solvers.
The article is concluded with future prospects in Section~\ref{sec:conclusion}.

@@

Solvers not complying with either of the above criteria were only evaluated in a so-called No-Limits track and were not eligible for the Main track awards.
The No-Limits thus enabled participation of closed-source solvers (not being able or willing to expose the source code for legal or other reasons) 
as well as portfolio solvers (combining two or more core SAT solvers developed by different groups of authors; c.f.~Sect.~\ref{sec:rules}).
However, solvers in No-Limits still competed against all other solvers submitted to the Main Track.
Thus, to deserve a mention, a No-Limits solver would need to beat all the Main Track participants.
The No-Limits track was only evaluated with respect to the benchmarks instances which were \emph{newly} submitted to SAT Competition 2020.

@@

Finally, the Main Track also had the so-called ``Glucose hack'' sub-track. 
Since in the past several advances in SAT solving required only a small modification of an established solver
to achieve a considerable contribution, this sub-track encouraged participation 
of small modifications of the Glucose 3.0 SAT solver. The limit for being considered a ``hack''
was set to 1000 non-space character edit distance from the sources of Glucose provided by the organizers. 
Unfortunately, in 2020 there were not enough participants in this sub-track and we do not report on it 
in the results section.

@@

In the Incremental Library track we mimic scenarios
where a SAT solver is used as a back-end solver in a more complex tool
(typically solving a harder problem than SAT) and is called multiple times before 
the enclosing tool reaches its final state. Incremental means that
the individual calls to the SAT solver are not independent, but may share 
a common subset of the input clauses or differ in the presence of additional 
unit clause assumptions~\cite{Nadel:2014:Incremental,Fazekas:2019:IncrementalInprocessing}.

Instead of using or extending the DIMACS input format, we realize the Incremental Library track
by relying on the general incremental interface IPASIR (Re-entrant Incremental Solver API) 
for SAT applications~\cite{Balyo:2015:SATRace}. The idea is that we actually run the 
enclosing tool on its own benchmark and communicate with the competing SAT solver 
through this interface API. Not only does the SAT solver in this track
need to solve a sequence of related problems fast, but its answers to the early questions
may in general influence which questions will follow next.
% E.g. Smaller unsat cores may allow the enclosing tool to converge using fewer SAT calls!

@@

The Parallel track evaluates solver performance in wall-clock time on multiple cores.
The benchmarks and the timeout (5000 seconds) are the same as in the Main track. 
Each solver run in this track was on a m4.16xlarge machine from Amazon Web Services (AWS).
These machines have 64 virtual cores and 256GB of memory. In contrast to the 
Main track, there is no proof logging for unsatisfiable benchmarks in the Parallel track.

@@

New this year was the Cloud track to evaluated distributed solvers. Similar to the
Parallel track, the solvers ran on the Amazon cloud, but using 100 machines of the
type m4.4xlarge, which have 16 virtual cores and 64GB of memory. 
Communication between the machines is possible using MPI and SSH.
The wall-clock timeout is 1000 seconds and the same benchmarks were used as in
the Main and Parallel track.

@@

Proofs of unsatisfiability were to be emitted in the DRAT format~\cite{DRATtrim},
either in its textual version, which is also very similar to the DIMACS input format,
or on a more compact binary version (for more details, see \cite{satComp2020www}, Unsat Certificates).
%
Details on certification are further discussed below in Section~\ref{sec:certif}.

@@

In all tracks it was required to output a model to certify recognizing a satisfiable instance.
On the other hand, certificates for unsatisfiable instances (proofs) were required only 
in the Main track (besides the No Limits track).
In some cases solvers output the correct result but the respective certificate was wrong. 
Such solvers were demoted to the No-limits track of the competition.

@@

Four solver participating in the Main track (\solver{Kissat-sat}, \solver{Relaxed-newTech}, \solver{CaDiCaL-alluip-trial}, and \solver{CMS-ccnr-lsids}) 
performed significantly better compared to the others, see Table~\ref{tab:results}.
Figure~\ref{main:all} shows the cumulative solved instances plot of the best performing solver of the strongest
10 teams (in short, the top 10 solvers)  together with the Virtual Best Solver (VBS).
The four top solvers all have a different code base, which is rare and has not been observed in many years.
Also, in recent competitions, the top solvers were much closer together.
The winning solver is \solver{Kissat-sat} and the runner-up is \solver{Relaxed-newTech}. Notice that 
\solver{Relaxed-newTech} solved more instances within the 2000 seconds limit.
The third place, based on the PAR-2 score, was for \solver{CMS-ccnr-lsids}. It solved two fewer instances
compared to \solver{CaDiCaL-alluip-trial}, but solved various formulas much more efficiently. 
Since the SAT Competition started to rank based on the PAR-2 score, such cases have been observed several times.

@@

The solver performances on unsatisfiable formulas was much closer,
see Figure~\ref{main:unsat}. Only \solver{Kissat-unsat}, the winner of the Main UNSAT track, 
performed significantly better compared to all other participants. It is therefore not surprising
that the VBS is reasonably close to \solver{Kissat-unsat}. \solver{Kissat} is a new solver that has been implemented in C, while most solvers are implemented in C++. Further 
details about \solver{Kissat} are described in Section~\ref{sec:kissat}.
The solvers \solver{CaDiCaL-trial}
and \solver{f2trc-s} won the second, respectively, third place in the Main UNSAT track.

\begin{figure}[ht]
\centering
\includegraphics[width=.9\textwidth]{img/paper-main-top10-UNSAT.pdf}
\caption{Performance of the top 10 solvers + VBS on Main track unsatisfiable benchmarks}
\end{figure}
\label{main:unsat}

The ranking in the Planning Track was closer. \solver{CaDiCaL-alliup-trial}, the winning solver, 
solved only one instance more compared to the runner up \solver{CMS-ccnr-lsids}. Their PAR-2 score is 
quite close as well. The third ranked solver, \solver{Kissat}, solved fewer instances, but fast runtime
on several instances resulted in a strong PAR-2 score. Notice that these solvers were also strong in the Main tracks. 
Apparently none of the participating solvers was optimized for planning instances.

The winning solver in the Parallel track is \solver{Painless-MCOMSPS-STR32}. The number in the post-fix denotes
that this solver used 32 threads on the 64 virtual cores that were available. It has been observed in earlier 
competitions that uses fewer threads than the number of virtual cores can be helpful as threads compete 
for memory, which can reduce of the overall performance. The runner up is \solver{Plingeling}, while the
third place is for \solver{ManyGlucose-32}. Figure~\ref{fig:res-parallel} shows the performance of
all participating parallel solvers. Notice in Table~\ref{tab:results} that only the \solver{Painless-MCOMSPS-STR} solvers
and \solver{Plingeling} had a lower PAR-2 score than the winner of the Main Track (\solver{Kissat-sat}). 
This shows that it is hard to beat sequential solvers with a parallel solver on the competition instances. 
We will show later in the article (Table~\ref{tab:tuples}) that small sequential portfolios 
perform better (in terms of PAR-2 score) than the parallel solvers.

@@

The clear winner of the Cloud track is \solver{Mallob-Mono} and the runner-up is \solver{TopoSAT2},
see Figure~\ref{fig:res-cloud}.
\solver{Mallob-Mono} was able to solver more instances in 1000 seconds, compared to the 
winner of the Parallel Track in 5000 seconds, thereby showing the potential of distributed SAT solving. 
The other four participants performed significantly worse. This is likely due to the initial challenges
to develop a distributed solver. This was the first year of this track. We expect that the gap will be 
closer is future years.

>>>>

\begin{abstract}
The SAT Competitions constitute a well-established series of yearly open international algorithm implementation competitions,
focusing on the Boolean satisfiability (or propositional satisfiability, SAT) problem. 
In this article, we provide
a detailed account on the 2020 instantiation of the SAT Competition, including the 
 new competition tracks and benchmark selection procedures, overview of solving strategies implemented in top-performing solvers, 
and a detailed analysis of the empirical data obtained from running the competition.
%SAT Competition 2020 stands in the tradition of the series of annual competitive events which motivate and assess the progress in SAT solving. 
%This competition was special as it introduced the new cloud track where SAT Solvers that run on hundreds of processors could compete. 
%Another novelty was the application-specific sub-track of the main track, 
%where solvers competed in solving benchmark instances from one specific domain only, and in this year that was the planning domain. 
%We used new tools to select and distribute benchmark instances and their attributes. 
%In this paper we provide a description of the well-known and the new competition tracks and how we organized them. 
%It follows then a detailed analysis of the results and the strategies of the award winning solvers. 
\end{abstract}

@@

The success of SAT would not be possible without the persistent efforts of the SAT community to further develop the performance and robustness
of SAT solvers. The SAT Competition series, with a history dating back to early 90s, aims to support and provide further incentives 
for maintaining and furthering this progress. Organized yearly as an international open event, the SAT competitions (and their variants
in the form of SAT Races and SAT Challenge) have a consistent track record of receiving tens of solver submissions yearly, submitted
by the community at large for obtaining a snapshot of the current state-of-the-art in practical SAT solving. Alongside participating solvers,
the competition invites through open calls submissions of benchmark instances representing, in particular, new interesting applications 
scenarios of SAT solvers. Indeed, in addition to evaluating recently developed solvers, an important aspect of the SAT competition series is
to collect on a yearly basis new benchmark sets, consisting of instances from various different application settings, which together with benchmark sets from previous years constitute a standard dataset for use in research papers and SAT solver development.

This article focuses on the 2020 instantiation of the SAT competitions. 
To this end, we provide
a detailed account of SAT Competition 2020 in terms of organizational details, competition tracks,
participating solvers, benchmarks, and the empirical results from the competition.
In terms of competition tracks, two new tracks, namely the cloud track and an application-specific track, where introduced in 2020,
in addition to the already earlier established main, parallel, and incremental tracks; we
provide motivation and the new organizational details for both of these new tracks.
In terms of solvers, we provide an overview of solving strategies and other details implemented in the top-performing solvers
from the competition, complementing the individual solver descriptions available in the 2020 competition proceedings \cite{SC2020}.
 As for benchmarks, we describe how the 2020 benchmark sets were constructed for each of the competition tracks, with 
an overview of the benchmarks contributed to the 2020 competition.
In terms of empirical results we provide further analysis on the results going beyond the standard rankings provided
on the SAT competition web pages.\footnote{\url{https://satcompetition.github.io/2020/}}
Finally, we also provide a discussion on lessons learned and ideas for future editions of SAT competitions.

The rest of this article is organized as follows. 
We start by providing an overview on the competition, including details on and motivations for the several
competition tracks, the rules and other technical requirements of the competition, the ranking schemes used
in evaluating the competing solvers, 
and the computing environments used for executing the competition
(Section~\ref{sec:overview}).
We then provide an overview of the benchmark sets used in evaluating the solvers, including their origins and the selection
process used for constructing the sets (Section~\ref{sec:instances}).
In Section~\ref{sec:results} we provide an overview of the competition results %together with a survey on the
followed by a survey on 
the solving strategies implemented distinctly in the top-ranking solvers in Section~\ref{sec:winners}. 
Going considerably beyond the plain competition rankings, in Section~\ref{sec:analysis}, we provide an in-depth analysis of the competition 
data from different perspectives, including correlation analysis of runtime performance of solvers and marginal contributions
of individual solvers to the ``virtual best solver'' and 
portfolios constructed from the competing solvers.
The article is concluded with future prospects in Section~\ref{sec:conclusion}.

@@

Solvers not complying with either of the above criteria were only evaluated in a so-called No-Limits track and were not eligible for the Main track awards.
The No-Limits track thus enabled participation of closed-source solvers (not being able or willing to expose the source code for legal or other reasons) 
as well as portfolio solvers (combining two or more core SAT solvers developed by different groups of authors; c.f.~Sect.~\ref{sec:rules}).
However, solvers in No-Limits still competed against all other solvers submitted to the Main Track.
Thus, to deserve a mention, a No-Limits solver would need to beat all the Main Track participants.
The No-Limits track was only evaluated with respect to the benchmarks instances which were \emph{newly} submitted to SAT Competition 2020.

@@

Finally, the Main Track also had the so-called ``Glucose hack'' sub-track. 
Since in the past several advances in SAT solving required only a small modification of an established solver
to achieve a considerable contribution, this sub-track encouraged participation 
of small modifications of the Glucose 3.0 SAT solver. The limit for being considered a ``hack''
was arbitrarily set to 1000 non-space character edit distance from the sources of Glucose provided by the organizers. 
Unfortunately, in 2020 there were not enough participants in this sub-track and we do not report on it 
in the results section.

@@

In the Incremental Library track we mimic scenarios
where a SAT solver is used as a back-end solver in a more complex tool
(typically solving a harder problem than SAT) and is called multiple times before 
the enclosing tool reaches its final state. ``Incremental'' means that
the individual calls to the SAT solver are not independent, but may share 
a common subset of the input clauses or differ in the presence of additional 
unit clause assumptions~\cite{Nadel:2014:Incremental,Fazekas:2019:IncrementalInprocessing}.

Instead of using or extending the DIMACS input format, we realize the Incremental Library track
by relying on the general incremental interface IPASIR (Re-entrant Incremental Solver API) 
for SAT applications~\cite{Balyo:2015:SATRace}. The idea is that we actually run the 
enclosing tool on its own benchmark and communicate with the competing SAT solver 
through this interface API. 
%
Not only does the SAT solver in this track
need to solve a sequence of related problems fast, but its answers to the early questions
may in general influence which questions will follow next.
% E.g. Smaller unsat cores may allow the enclosing tool to converge using fewer SAT calls!

@@

The Parallel track evaluates solver performance in wall-clock time on multiple cores.
The benchmarks and the timeout (5000 seconds) are the same as in the Main track. 
Each solver in this track was run on a m4.16xlarge machine from Amazon Web Services (AWS).
These machines have 64 virtual cores and 256GB of memory. In contrast to the 
Main track, proof logging for unsatisfiable instances is not required in the Parallel track.

@@

New this year was the Cloud track, dedicated to evaluating distributed solvers. Similar to the
Parallel track, the solvers ran on the Amazon cloud, but using 100 machines of the
type m4.4xlarge, which have 16 virtual cores and 64GB of memory. 
Communication between the machines is possible using MPI and SSH.
The wall-clock timeout is 1000 seconds and the same benchmarks were used as in
the Main and Parallel track.

@@

Proofs of unsatisfiability were to be emitted in the DRAT format~\cite{DRATtrim},
either in its textual version, which is also very similar to the DIMACS input format,
or in a more compact binary version (for more details, see \cite{satComp2020www}, Unsat Certificates).
%
Details on certification are further discussed below in Section~\ref{sec:certif}.

@@

In all tracks it was required to output a model to certify recognizing a satisfiable instance.
On the other hand, certificates for unsatisfiable instances (proofs) were required only 
in the Main track (besides the No Limits track).
In some cases, a solver output the correct result, but the respective certificate was wrong. 
Such solvers were demoted to the No-limits track of the competition.

@@

Four solvers participating in the Main track (\solver{Kissat-sat}, \solver{Relaxed-newTech}, \solver{CaDiCaL-alluip-trial}, and \solver{CMS-ccnr-lsids}) 
performed significantly better compared to the others, see Table~\ref{tab:results}.
Figure~\ref{main:all} shows the cumulative solved instances plot of the best performing solver of the strongest
10 teams (in short, the top 10 solvers)  together with the Virtual Best Solver (VBS).
The four top solvers all have a different code base, which is rare and has not been observed for many years.
Also, in recent competitions, the top solvers were much closer together.
The winning solver is \solver{Kissat-sat} and the runner-up is \solver{Relaxed-newTech}. Notice that 
\solver{Relaxed-newTech} solved more instances within the 2000 seconds limit.
The third place, based on the PAR-2 score, was for \solver{CMS-ccnr-lsids}. It solved two fewer instances
compared to \solver{CaDiCaL-alluip-trial}, but solved various formulas much more quickly. 
Since the SAT Competition started to rank based on the PAR-2 score, such cases have been observed several times.

@@

The solver performance on unsatisfiable formulas was much closer,
see Figure~\ref{main:unsat}. Only \solver{Kissat-unsat}, the winner of the Main UNSAT track, 
performed significantly better compared to all other participants. It is therefore not surprising
that the VBS is reasonably close to \solver{Kissat-unsat}. \solver{Kissat} is a new solver that has been implemented in C, while most solvers are implemented in C++. Further 
details about \solver{Kissat} are given in Section~\ref{sec:kissat}.
The solvers \solver{CaDiCaL-trial}
and \solver{f2trc-s} placed, respectively, second and third in the Main UNSAT track.

\begin{figure}[ht]
\centering
\includegraphics[width=.9\textwidth]{img/paper-main-top10-UNSAT.pdf}
\caption{Performance of the top 10 solvers + VBS on Main track unsatisfiable benchmarks}
\label{main:unsat}
\end{figure}

The ranking in the Planning Track was closer. \solver{CaDiCaL-alliup-trial}, the winning solver, 
solved only one instance more compared to the runner up \solver{CMS-ccnr-lsids}. Their PAR-2 score is 
quite close as well. The third ranked solver, \solver{Kissat}, solved fewer instances, but its fast runtime
on several instances resulted in a strong PAR-2 score. Notice that these solvers were also strong in the Main track. 
Apparently none of the participating solvers was optimized for planning instances.

The winning solver in the Parallel track is \solver{Painless-MCOMSPS-STR32}. The number in the suffix denotes
that this solver used 32 threads on the 64 virtual cores that were available. It has been observed in earlier 
competitions that using fewer threads than the number of virtual cores can be helpful as threads compete 
for memory, which can reduce of the overall performance. The runner up is \solver{Plingeling}, while the
third place is taken by \solver{ManyGlucose-32}. Figure~\ref{fig:res-parallel} shows the performance of
all participating parallel solvers. Notice in Table~\ref{tab:results} that only the \solver{Painless-MCOMSPS-STR} solvers
and \solver{Plingeling} had a lower PAR-2 score than the winner of the Main Track (\solver{Kissat-sat}). 
This shows that it is hard to beat sequential solvers with a parallel solver on the competition instances. 
We will show later in the article (Table~\ref{tab:tuples}) that small sequential portfolios 
perform better (in terms of PAR-2 score) than the parallel solvers.

@@

The clear winner of the Cloud track is \solver{Mallob-Mono} and the runner-up is \solver{TopoSAT2}
(see Figure~\ref{fig:res-cloud}).
\solver{Mallob-Mono} was able to solver more instances in 1000 seconds than the 
winner of the Parallel Track in 5000 seconds, thereby showing the potential of distributed SAT solving. 
The other four participants performed significantly worse. This is likely due to the initial challenges
in developing a distributed solver. 2020 was the first year of this track. We expect that the gap will be 
smaller in the future years.
